Titanic 
========================================================
## R packages
```{r}
library(dplyr)
library(stringr)
library(randomForest)
library(glmnet)
```
## Data input
```{r}
df <- read.csv('train.csv',stringsAsFactors=FALSE)
df.test <- read.csv('test.csv',stringsAsFactors=FALSE)
```
We check the summary of data, and the ratio of missing values.
```{r}
str(df)
summary(df)

options(digits=3)

rbind(na = colMeans(is.na(df)), nullstr = colMeans(df == ""))
rbind(na = colMeans(is.na(df.test)), nullstr = colMeans(df.test == ""))
```

## Feature Engineering
A few features are created or modified.  For cabin, we get the first capital letter.  For title, we extract the middle pattern of the name field.  
```{r}

# extract Cabin/Embarked/Title
extractFeatures <- function(df) {
  df %.% 
    mutate(Cabin = str_extract(Cabin,'^[A-Z]')) %.%
    mutate(Embarked = str_extract(Embarked,'^[CQS]')) %.%
    mutate(Title = str_extract(str_extract(Name, ', \\w+\\.'), 
                               '\\w+'))    
}
```
A few features are removed.
```{r}
removeFeatures <- function(df) {
  df %.% select(-Ticket, -Name, -PassengerId)
}

```

```{r}
calculateImputeValues <- function(df) {
  lapply(df, function(col) 
    ifelse(is.numeric(col), mean(col, na.rm=T), "Unknown"))
}
imputeFeatures <- function(df, impute.values) {
  for (c in names(df)) {
      df[[c]] <- ifelse(is.na(df[[c]]), impute.values[[c]], df[[c]])
  }
  df
}

```

```{r}
chr.features.to.factor <- function(df, use.ref=F, df.ref=NA) {
  for (c in names(df)) {
    if (is.character(df[[c]])) {
      if (!use.ref) {
        df[[c]] <- factor(df[[c]])
      }  
      else {
        df[[c]] <- factor(df[[c]], levels=levels(df.ref[[c]]))
      }  
      df[[c]][is.na(df[[c]])] <- "Unknown"
    }
  }
  df  
}
  
df2 <- extractFeatures(df)
df2 <- removeFeatures(df2)
impute.train <- calculateImputeValues(df2)
df2 <- imputeFeatures(df2, impute.train)
df2 <- chr.features.to.factor(df2)
df2 <- df2 %.% mutate(Survived = factor(Survived))

```
Simliar processing are applied to test set.
```{r}
df.test2 <- extractFeatures(df.test)
df.test2 <- removeFeatures(df.test2)

df.test2 <- imputeFeatures(df.test2, impute.train)
df.test2 <- chr.features.to.factor(df.test2, use.ref=T, df2)
```
## Random forest model
```{r}
seed = 123
set.seed(seed)
fit.rf <- randomForest(Survived ~  ., data = df2)
fit.rf

pred <- predict(fit.rf,newdata=df.test2)
df.out <- data.frame(PassengerId = df.test$PassengerId, 
                     Survived = pred)
write.csv(df.out, file="randomForest_1.csv", 
          row.names=F, col.names=T, quote=F)
```

### Logistic regression model
```{r}

set.seed(seed)
fit.glm <- cv.glmnet(data.matrix(df2 %.% select(-Survived)), df2$Survived,  alpha=1, family="binomial")

coef(fit.glm)
pred <- predict(fit.glm, newx=data.matrix(df.test2), s = "lambda.min", type="class")
df.out <- data.frame(PassengerId = df.test$PassengerId, 
                     Survived = as.vector(pred))
write.csv(df.out, file="glmnet_1.csv", 
          row.names=F, col.names=T, quote=F)

```

